---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Admin work: load required packages
library(readr)
library(dplyr)
library(data.table)
library(ggplot2)
library(lubridate)
library(mice)
library(lattice)

#Directory for project (assumes project repo has been downloaded to working directory)

projectfile<-"./activity.csv"

```


## Loading and preprocessing the data
  
Read in and clean the data
```{r}


activitydata<-read.csv(projectfile,colClasses = c("integer","character","integer"))


activitydata$date<-as.Date(activitydata$date)
activitydata<-mutate(activitydata,weekday=wday(date))
activitydata<-arrange(activitydata,interval)
```
## What is mean total number of steps taken per day?
```{r}
daysums<-tapply(activitydata$steps,activitydata$date,mean,na.rm=T)
hist(daysums)
stepaverage<-mean(daysums,na.rm = T)
stepmedian<-median(daysums,na.rm = T)
print(stepaverage)
print(stepmedian)
```
Mean of steps taken each day is `r stepaverage `, and median is `r stepmedian`.  

## What is the average daily activity pattern?
```{r}
groupedactivitydata<-group_by(activitydata,interval)
groupedactivitydata<-mutate(groupedactivitydata,intervalaverage=mean(steps,na.rm = T))
with(groupedactivitydata,plot(interval,intervalaverage,xlab="Time",type="l"))
dailymax<-max(groupedactivitydata$intervalaverage)
maxrows<-which(groupedactivitydata$intervalaverage==dailymax)
maxinterval<-unique(groupedactivitydata$interval[maxrows])
print(maxinterval)
```

## Imputing missing values
  
  
Using the "mice" impute function from the "mice package
```{r results='hide'}
imputed_Data <- mice(groupedactivitydata, m=1, maxit = 10, method = 'pmm', seed = 500)
completeData <- complete(imputed_Data)
```
```{r}
imputeddaysums<-tapply(completeData$steps,completeData$date,mean)
hist(imputeddaysums)
print(summary(imputeddaysums))

```
Comparing to the previous histogram, median, and mean, we see that imputing the NAs does not make a significant difference.  

## Are there differences in activity patterns between weekdays and weekends?
```{r}
completeData<-mutate(completeData,weekdaytype=factor(completeData$weekday,                                                    levels=1:7,labels=c("weekend","weekday","weekday","weekday"                                               ,"weekday","weekday","weekend")))
completeData<-group_by(completeData,weekdaytype,interval)
completeData<-mutate(completeData,intervalaverage=mean(steps,na.rm = T))
xyplot(intervalaverage~interval|weekdaytype,completeData,type="l",layout=c(1,2))
```
  
  
  
As we can see, weekdays have a strong spike in activity around 0830, whereas on weekends the activity is more
evenly distributed across the waking hours.